---
title: 'P&S-2022: Lab assignment 2'
author: "Name1, Name2, Name3"
output:
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

## General comments and instructions

-   Complete solution will give you **4 points** (working code with
    explanations + oral defense). Submission deadline **November 1,
    2023, 22:00**\
-   The report must be prepared as an *R notebook*; you must submit to
    **cms** both the source *R notebook* **and** the generated html
    file\
-   At the beginning of the notebook, provide a work-breakdown structure
    estimating efforts of each team member\
-   For each task, include
    -   problem formulation and discussion (what is a reasonable answer
        to discuss);\
    -   the corresponding $\mathbf{R}$ code with comments (usually it is
        just a couple of lines long);\
    -   the statistics obtained (like sample mean or anything else you
        use to complete the task) as well as histograms etc to
        illustrate your findings;\
    -   justification of your solution (e.g. refer to the corresponding
        theorems from probability theory);\
    -   conclusions (e.g. how reliable your answer is, does it agree
        with common sense expectations etc)\
-   The **team id number** referred to in tasks is the **two-digit**
    ordinal number of your team on the list. Include the line
    **set.seed(team id number)** at the beginning of your code to make
    your calculations reproducible. Also observe that the answers **do**
    depend on this number!\
-   Take into account that not complying with these instructions may
    result in point deduction regardless of whether or not your
    implementation is correct.

### Task 1

#### In this task, we discuss the $[7,4]$ Hamming code and investigate its reliability. That coding system can correct single errors in the transmission of $4$-bit messages and proceeds as follows:

-   given a message $\mathbf{m} = (a_1 a_2 a_3 a_4)$, we first encode it
    to a $7$-bit *codeword*
    $\mathbf{c} = \mathbf{m}G = (x_1 x_2 x_3 x_4 x_5 x_6 x_7)$, where
    $G$ is a $4\times 7$ *generator* matrix\
-   the codeword $\mathbf{c}$ is transmitted, and $\mathbf{r}$ is the
    received message\
-   $\mathbf{r}$ is checked for errors by calculating the *syndrome
    vector* $\mathbf{z} := \mathbf{r} H$, for a $7 \times 3$
    *parity-check* matrix $H$\
-   if a single error has occurred in $\mathbf{r}$, then the binary
    $\mathbf{z} = (z_1 z_2 z_3)$ identifies the wrong bit no.
    $z_1 + 2 z_2 + 4z_3$; thus $(0 0 0)$ shows there was no error (or
    more than one), while $(1 1 0 )$ means the third bit (or more than
    one) got corrupted\
-   if the error was identified, then we flip the corresponding bit in
    $\mathbf{r}$ to get the corrected
    $\mathbf{r}^* = (r_1 r_2 r_3 r_4 r_5 r_6 r_7)$;\
-   the decoded message is then $\mathbf{m}^*:= (r_3r_5r_6r_7)$.

#### The **generator** matrix $G$ and the **parity-check** matrix $H$ are given by

$$  
    G := 
    \begin{pmatrix}
        1 & 1 & 1 & 0 & 0 & 0 & 0 \\
        1 & 0 & 0 & 1 & 1 & 0 & 0 \\
        0 & 1 & 0 & 1 & 0 & 1 & 0 \\
        1 & 1 & 0 & 1 & 0 & 0 & 1 \\
    \end{pmatrix},
 \qquad 
    H^\top := \begin{pmatrix}
        1 & 0 & 1 & 0 & 1 & 0 & 1 \\
        0 & 1 & 1 & 0 & 0 & 1 & 1 \\
        0 & 0 & 0 & 1 & 1 & 1 & 1
    \end{pmatrix}
$$

#### Assume that each bit in the transmission $\mathbf{c} \mapsto \mathbf{r}$ gets corrupted independently of the others with probability $p = \mathtt{id}/100$, where $\mathtt{id}$ is your team number. Your task is the following one.

1.  Simulate the encoding-transmission-decoding process $N$ times and
    find the estimate $\hat p$ of the probability $p^*$ of correct
    transmission of a single message $\mathbf{m}$. Comment why, for
    large $N$, $\hat p$ is expected to be close to $p^*$.\
2.  By estimating the standard deviation of the corresponding indicator
    of success by the standard error of your sample and using the CLT,
    predict the \emph{confidence} interval
    $(p^*-\varepsilon, p^* + \varepsilon)$, in which the estimate
    $\hat p$ falls with probability at least $0.95$.\
3.  What choice of $N$ guarantees that $\varepsilon \le 0.03$?\
4.  Draw the histogram of the number $k = 0,1,2,3,4$ of errors while
    transmitting a $4$-digit binary message. Do you think it is one of
    the known distributions?

#### You can (but do not have to) use the chunks we prepared for you

#### First, we set the **id** of the team and define the probability $p$ and the generator and parity-check matrices $G$ and $H$

```{r}
# your team id number 
                          ###
id <- 50                  ### Change to the correct id!
                          ###
set.seed(id)
p <- id/100
# matrices G and H
G <- matrix(c(1, 1, 1, 0, 0, 0, 0,
		1, 0, 0, 1, 1, 0, 0,
		0, 1, 0, 1, 0, 1, 0,
		1, 1, 0, 1, 0, 0, 1), nrow = 4, byrow = TRUE)
H <- t(matrix(c(1, 0, 1, 0, 1, 0, 1,
		0, 1, 1, 0, 0, 1, 1,
		0, 0, 0, 1, 1, 1, 1), nrow = 3, byrow = TRUE))
# cat("The matrix G is: \n") 
#G  
#cat("The matrix H is: \n") 
#H
#cat("The product GH must be zero: \n")
#(G%*%H) %%2
```

#### Next, generate the messages

```{r}
# generate N messages

message_generator <- function(N) {
  matrix(sample(c(0,1), 4*N, replace = TRUE), nrow = N)
}  
messages <- message_generator(100)
codewords <- (messages %*% G) %% 2
```

#### Generate random errors; do not forget that they occur with probability $p$! Next, generate the received messages

```{r}
#  errors <- ...
#  received <-   
```

The next steps include detecting the errors in the received messages,
correcting them, and then decoding the obtained messages. After this,
you can continue with calculating all the quantities of interest

**Do not forget to include several sentences summarizing your work and
the conclusions you have made!**

### Task 2.

#### In this task, we discuss a real-life process that is well modelled by a Poisson distribution. As you remember, a Poisson random variable describes occurrences of rare events, i.e., counts the number of successes in a large number of independent random experiments. One of the typical examples is the **radioactive decay** process.

#### Consider a sample of radioactive element of mass $m$, which has a big *half-life period* $T$; it is vitally important to know the probability that during a one second period, the number of nuclei decays will not exceed some critical level $k$. This probability can easily be estimated using the fact that, given the *activity* ${\lambda}$ of the element (i.e., the probability that exactly one nucleus decays in one second) and the number $N$ of atoms in the sample, the random number of decays within a second is well modelled by Poisson distribution with parameter $\mu:=N\lambda$. Next, for the sample of mass $m$, the number of atoms is $N = \frac{m}{M} N_A$, where $N_A = 6 \times 10^{23}$ is the Avogadro constant, and $M$ is the molar (atomic) mass of the element. The activity of the element, $\lambda$, is $\log(2)/T$, where $T$ is measured in seconds.

#### Assume that a medical laboratory receives $n$ samples of radioactive element ${{}^{137}}\mathtt{Cs}$ (used in radiotherapy) with half-life period $T = 30.1$ years and mass $m = \mathtt{team\, id \,number} \times 10^{-6}$ g each. Denote by $X_1,X_2,\dots,X_n$ the **i.i.d. r.v.**'s counting the number of decays in sample $i$ in one second.

1.  Specify the parameter of the Poisson distribution of $X_i$ (you'll
    need the atomic mass of *Cesium-137*)\
2.  Show that the distribution of the sample means of $X_1,\dots,X_n$
    gets very close to a normal one as $n$ becomes large and identify
    that normal distribution. To this end,
    -   simulate the realization $x_1,x_2,\dots,x_n$ of the $X_i$ and
        calculate the sample mean $s=\overline{\mathbf{x}}$;
    -   repeat this $K$ times to get the sample
        $\mathbf{s}=(s_1,\dots,s_K)$ of means and form the empirical
        cumulative distribution function $\hat F_{\mathbf{s}}$ of
        $\mathbf{s}$;
    -   identify $\mu$ and $\sigma^2$ such that the \textbf{c.d.f.} $F$
        of $\mathscr{N}(\mu,\sigma^2)$ is close to the \textbf{e.c.d.f.}
        $\hat F_{\mathbf{s}}$ and plot both **c.d.f.**'s on one graph to
        visualize their proximity (use the proper scales!);
    -   calculate the maximal difference between the two
        \textbf{c.d.f.}'s;
    -   consider cases $n = 5$, $n = 10$, $n=50$ and comment on the
        results.\
3.  Calculate the largest possible value of $n$, for which the total
    number of decays in one second is less than $8 \times 10^8$ with
    probability at least $0.95$. To this end,
    -   obtain the theoretical bound on $n$ using Markov inequality,
        Chernoff bound and Central Limit Theorem, and compare the
        results;\
    -   simulate the realization $x_1,x_2,\dots,x_n$ of the $X_i$ and
        calculate the sum $s=x_1 + \cdots +x_n$;
    -   repeat this $K$ times to get the sample
        $\mathbf{s}=(s_1,\dots,s_K)$ of sums;
    -   calculate the number of elements of the sample which are less
        than critical value ($8 \times 10^8$) and calculate the
        empirical probability; comment whether it is close to the
        desired level $0.95$

```{r}
lambda <- 1  # change this!
N <- 100     # change this!
mu <- N * lambda
K <- 1e3
n <- 5
sample_means <- colMeans(matrix(rpois(n*K, lambda = mu), nrow=n))
```

#### Next, calculate the parameters of the standard normal approximation

```{r}
mu <- 0       # change this!
sigma <- 1    # change this!
```

#### We can now plot ecdf and cdf

```{r}
xlims <- c(mu-3*sigma,mu+3*sigma)
Fs <- ecdf(sample_means)
plot(Fs, 
     xlim = xlims, 
     ylim = c(0,1),
     col = "blue",
     lwd = 2,
     main = "Comparison of ecdf and cdf")
curve(pnorm(x, mean = mu, sd = sigma), col = "red", lwd = 2, add = TRUE)
```

**Next, proceed with all the remaining steps**

**Do not forget to include several sentences summarizing your work and
the conclusions you have made!**

### Task 3.

#### In this task, we use the Central Limit Theorem approximation for continuous random variables.

#### One of the devices to measure radioactivity level at a given location is the Geiger counter. When the radioactive level is almost constant, the time between two consecutive clicks of the Geiger counter is an exponentially distributed random variable with parameter $\nu_1 = \mathtt{team\,id\,number} + 10$. Denote by $X_k$ the random time between the $(k-1)^{\mathrm{st}}$ and $k^{\mathrm{th}}$ click of the counter.

1.  Show that the distribution of the sample means of
    $X_1, X_2,\dots,X_n$ gets very close to a normal one (which one?) as
    $n$ becomes large. To this end,
    -   simulate the realizations $x_1,x_2,\dots,x_n$ of the
        \textbf{r.v.} $X_i$ and calculate the sample mean
        $s=\overline{\mathbf{x}}$;\
    -   repeat this $K$ times to get the sample
        $\mathbf{s}=(s_1,\dots,s_K)$ of means and then the
        \emph{empirical cumulative distribution} function
        $F_{\mathbf{s}}$ of $\mathbf{s}$;\
    -   identify $\mu$ and $\sigma^2$ such that the \textbf{c.d.f.} of
        $\mathscr{N}(\mu,\sigma^2)$ is close to the \textbf{e.c.d.f.}
        $F_{\mathbf{s}}$ of and plot both \textbf{c.d.f.}'s on one graph
        to visualize their proximity;\
    -   calculate the maximal difference between the two
        \textbf{c.d.f.}'s;\
    -   consider cases $n = 5$, $n = 10$, $n=50$ and comment on the
        results.
2.  The place can be considered safe when the number of clicks in one
    minute does not exceed $100$. It is known that the parameter $\nu$
    of the resulting exponential distribution is proportional to the
    number $N$ of the radioactive samples, i.e., $\nu = \nu_1*N$, where
    $\nu_1$ is the parameter for one sample. Determine the maximal
    number of radioactive samples that can be stored in that place so
    that, with probability $0.95$, the place is identified as safe. To
    do this,
    -   express the event of interest in terms of the \textbf{r.v.}
        $S:= X_1 + \cdots + X_{100}$;\
    -   obtain the theoretical bounds on $N$ using the Markov
        inequality, Chernoff bound and Central Limit Theorem and compare
        the results;\
    -   with the predicted $N$ and thus $\nu$, simulate the realization
        $x_1,x_2,\dots,x_{100}$ of the $X_i$ and of the sum
        $S = X_1 + \cdots + X_{100}$;\
    -   repeat this $K$ times to get the sample
        $\mathbf{s}=(s_1,\dots,s_K)$ of total times until the
        $100^{\mathrm{th}}$ click;\
    -   estimate the probability that the location is identified as safe
        and compare to the desired level $0.95$

## Part 3.1

#### First, generate samples and sample means

```{r}
# Calculation of the parameter of exponential r.v.
nu1 <- 15 # 10 + id = 10 + 5

K <- 1e3

n <- c(5, 10, 50)

# Simulation of random variables and calculation of sample means for n=5
# Repeating this K times for n=5
sample_means <- colMeans(matrix(rexp(5*K, rate = nu1), nrow=5))
```

#### Next, calculate the parameters of the standard normal approximation

```{r}
mu <- 1 / nu1
sigma <- mu / sqrt(n)
```

#### We can now plot ecdf and cdf for n = 5, 10, 50

```{r}
# Creating a sequence of values from 0 to 0.500 with a step of 0.025 
# Then use this for finding difference between ECDF and CDF
x <- seq(0, 0.500, by=0.025)

for (i in n) {
  
  # Simulation of random variables and calculation of sample means for different n
  # Creation of a matrix and calculation of column means
  sample_means <- colMeans(matrix(rexp(i*K, rate = nu1), nrow=i))
  
  # Calculation of μ and σ^2
  mu <- 1 / nu1
  sigma <- mu / sqrt(i)
  
  # Setting the x-axis limits
  # Left and right bounds
  xlims <- c(mu-3*sigma,mu+3*sigma)
  
  # Empirical Cumulative Distribution Function (ECDF) calculation
  Fs <- ecdf(sample_means)
  
  # Visualizing ECDF
  plot(Fs, 
       xlim = xlims, 
       col = "blue",
       lwd = 2,
       main = paste("CDF and ECDF for n = ", i))
  
  # Visualize CDF
  curve(pnorm(x, mean = mu, sd = sigma), col = "red", lwd = 2, add = TRUE)
  
  # Calculating difference between CDF and ECDF
  print(max(abs(ecdf(sample_means)(x)-pnorm(x, mean = mu, sd = sigma))))
}
```
## Conclusion

According to the graphs, we can observe that with bigger n (in our case it`s 50) our two lines for ECDF and CDF overlap. Also as we can see the difference between this two function is the smallest for larger values of n.

## Part 3.2

#### First, consider the event of interest

$$S = X_1 + ... + X_{100}$$
As from the condition we know that number of clicks doesn`t exceed 100, so $n = 100$

```{r}
n = 100
S <- sum(rexp(n, rate = nu1))
S
```

#### Second, obtaining theoretical bounds

**Markov**

### Formula

$$P(S \ge a) \le \frac {\mu_S}{a}$$ 
It is known that the place can be considered safe when the number of clicks in one minute does not exceed
100, so we need to find probability, where $a=1$ (1 minute)

$$P(S \ge 1) \le \frac {\mu_S}{1}$$ 
Our ${\mu_S}$ is equal to:

$${\mu_S} = \frac {100}{\nu} \Rightarrow {\mu_S} = \frac {100}{15*N}$$
Now, we can substitute our values into inequality

$$P(S \ge 1) \le \frac {100} {15*N}$$ 
As given in the condition, the probability that our place is safe is 0,95 
Using this information we can find upper bound of Markov`s inequality

$$0.95 \le \frac {100} {15*N}$$ 
From this inequality we need to find $N$, which is he number N of the radioactive samples

$$N \le 7.02 $$ 

**Chernoff**

### Formula

$$P(S \ge a) \le e^{-ta}*M_S(t)\\t>0$$ 
We need to find $P(S\ge 1)$

$$P(S \ge 1) \le e^{-t}*M_S(t)$$ 

Now we need to find moment generating function (mgf)
$$M_S(t) = E(e^{St}) = E(e^{X_1t} * \dots * e^{X_{100}t}) = E^{100}(e^{X_{i}t}) = M_X^{100}(t) = (\frac{\nu_1*N} {\nu_1*N -  t})^{100}$$
After that we can limit our Chernoff 

$$P(S \ge 1) \le e^{-t}*M_s(t) \ge 0.95$$ 

Now we need to find $N$ from this inequality

$$N \le \frac{t*(e*0.95)^{1/100}}{15*((e*0.95)^{1/100} - 1)}$$
To find N we need to take derivative

$$N \le 6.06$$


**Central Limit theorem**

### Formula

$$P(\frac {S - 100 * \mu} {\sigma * \sqrt n} \le t) \rightarrow Ф(t)$$
We need to find 

$$P(S \ge 1) = 1 - P(S \le 1)$$
And we can limit our probability

$$1 - P(S \le 1) \ge 0.95 \Rightarrow P(S \le 1) \le 0.05$$
Now we find our ${\mu_s}$ and ${\sigma_s}$

$$\mu = \frac {1} {15*N} = \sigma$$ 
Now, let`s substitute our values to the inequality 
P.S. ${\sqrt n} = {\sqrt 100} = {10}$

$$P(\frac {S - 100 * \frac 1 {15 * N}}{ \frac 1 {15 * N} * 10} \le \frac {1 - 100 * \frac 1 {15 * N}}{\frac 1 {15 * N}* 10}) = Ф(\frac {1 - 100 * \frac 1 {15 * N}}{\frac 1 {15 * N}* 10})$$
$$P(Z \le \frac {1 - 100 * \frac 1 {15 * N}}{\frac 1 {15 * N}* 10}) \le 0.05$$
Now, we need to use z-score and z-table to find value of $N$
P.S. According to z-table the equivalent of 0.05 is -1.65

$$\frac {1 - 100 * {\mu }}{10 * {\sigma}} \le -1.65$$
$$\frac {1 - 100 * \frac {1} {15*N}}{10 * \frac {1} {15*N}} \le -1.65$$
Now, we calculate our $N$

$$N \le \frac {10 - 1.65}{0.1 * 15}$$
$$N \le 5.56$$

**Simulate the realization**

```{r}
# Lowest and highest value of N
lowest <- 5.56
highest <- 7.02

# Function for estimating new probabilities for values of N
estimation <- function(N, K, nu1) {
  nu <- N * nu1
  
  # Generating K samples of size n from an exponential distribution with rate parameter nu
  values <- replicate(K, sum(rexp(100, rate = nu)))
  
  # Check how many values are >= 1
  # Find probability of having a sum >= 1 for these samples
  prob <- sum(values >= 1) / K
  return(prob)
}

# Create a sequence of values from lowest to highest 
# This is done to cover the range of N values for which probabilities need to be estimated
val <- seq(lowest, highest, by = 0.001)

# Apply the estimation function to each value in the val sequence 
# This calculates the probability of having a sum greater than or equal to 1 for different values of N within the specified range
probs <- sapply(val, function(N) estimation(N, K, nu1))

cat("Probability for N = 5.56 =>", estimation(5.56, K, nu1), "\n")
cat("Probability for N = 6.06 =>", estimation(6.06, K, nu1), "\n")
cat("Probability for N = 7.02 =>", estimation(7.02, K, nu1), "\n")
```

## Conclusion

As we can observe, the most accurate result is obtained by using CLT because while comparing estimating probability with that that was given in the condition it`s the same. Markov`s inequality shows the worst result, while comparing it has the difference more than 0.5. Charnoff method of finding bound is closer to the right ansear but not as accurate as CLT.